---
title: "Portfolio_5"
author: "Lindley Slipetz"
date: "3/2/2021"
output: html_document
---

This project is looking at the representation of women in Philosophy through being professors and publishing in top journals. I have two data sets. One is the breakdown of professorship by gender at the top university for Philosophy in the US. Another is a breakdown of article authorship by gender. Unfortunately, I'm going to have to leave the non-binary data from the professor data out of the analysis because the publication dataset only includes binary genders.

We'll start by loading in packages.

```{r packages, warning = FALSE}
library(tidyverse)
#install.packages("pdftools")
library(pdftools)
library(readxl)
```

Now, let's load the data. For our datasets, we have an excel spreadsheet and a PDF...so this should be fun.

```{r load_data}
faculty <- read_excel("G:\\My Drive\\DataScience\\Portfolio_5\\Portfolio_5\\data\\faculty.xlsx", sheet = 1)
publishing <- pdf_text("G:\\My Drive\\DataScience\\Portfolio_5\\Portfolio_5\\data\\publishing.pdf")
```

